\documentclass[twoside]{article}
\input{settings.tex}

\begin{document}
\lecture{16}{Graphical Network Inference}{}{Sai Zhang}{}{The note is built on Prof. \hyperlink{http://faculty.marshall.usc.edu/jinchi-lv/}{Jinchi Lv}'s lectures of the course at USC, DSO 607, High-Dimensional Statistics and Big Data Problems.}
%\footnotetext{These notes are partially based on those of Nigel Mansell.}

\section{Motivation}
Consider a classic question: For $n$ observations of dimension $p$, how can we capture the statistical relationships between the variables of interest? Consider the example of the multivariate Gaussian distribution:
\begin{example}{Multivariate Gaussian Distribution}{}
    Suppose we have $n$ observations of dimension $p$, $\mathbf{x}\sim \mathcal{N}(\mathbf{0},\boldsymbol{\Sigma})$. let $\mathbf{S}$ be the empirical covariance matrix. Then the probability density 
    $$
    f(\mathbf{x}) = \frac{1}{(2\pi)^{p/2}\det (\boldsymbol{\Sigma})^{1/2}}\exp\left\{ -\frac{1}{2}(\mathbf{x}-\boldsymbol{\mu})'\boldsymbol{\Sigma}^{-1}(\mathbf{x}-\boldsymbol{\mu}) \right\}
    $$
    define the \textbf{inverse covariance matrix} or \textbf{precision matrix} as $ \boldsymbol{\Omega}=\boldsymbol{\Sigma}^{-1} $, then we have 
    $$
    f_{\mathbf{\mu},\boldsymbol{\Omega}} = \exp \left\{ \boldsymbol{\mu'\Omega x} - \left< \boldsymbol{\Omega},\frac{1}{2}\mathbf{xx}' \right> -\frac{p}{2}\log(2\pi) + \frac{1}{2} \log\det(\boldsymbol{\Omega}) - \frac{1}{2}\boldsymbol{\mu'\Omega\mu} \right\}
    $$
\end{example}
In this example, we know that \textbf{every} multivariate Gaussian distribution can be represented by a pairwise \myhl[myblue]{\textbf{Gaussian Markov Random Field (GMRF)}}, which an \textbf{\underline{undirected graph}} $G=(V,E)$
\begin{itemize}
    \item representing the collection of variables $\mathbf{x}$ by a vertex set $\mathcal{V}=\left\{1,\cdots,p\right\}$
    \item encoding correlations between variables by a set of edges $\mathcal{E}=\left\{ (i,j)\in \mathcal{V}\mid i=\neq j,\Omega_{ij}\neq 0 \right\}$  
\end{itemize}

\paragraph*{What's the goal?} We want to estimate a \myhl[myblue]{\textbf{sparse}} graph structure given $n\ll p$ i.i.d. observations. But what does sparsity means in this context? A sparse graph is \textbf{\underline{equivalent}} to a sparse precision matrix.

%\newpage
%\bibliographystyle{plainnat}
%\bibliography{ref.bib}

\end{document}